10th International Satisfiability Modulo Theories Competition (SMT-COMP 2015): Rules and Procedures

### 2015-02-17 rules/rules15.tex
First draft of 2015 rules added.

\documentclass[12pt]{article}
\usepackage{times}
\usepackage{fullpage}
\usepackage{url}
\usepackage{epsf}

@@

@@

\title{DRAFT\\Satisfiability Modulo Theories Competition (SMT-COMP) 2014: Rules and 
Procedures}

@@

\author{
David R. Cok \\
GrammaTech, Inc. \\
Ithaca, NY (USA) \\
cok@frontiernet.net \\
\and
David Deharbe \\
Federal University of Rio Grande do Norte \\
Brazil \\
David.Deharbe@loria.fr \\
\and
Tjark Weber \\
Uppsala University \\
Sweden \\
tjark.weber@it.uu.se \\ 
}

@@

\def\eg{\textit{e.g.}}
\def\ie{\textit{i.e.}}

\noindent Comments on this document should be emailed to the smtcomp mailing
list or, if necessary, directly to the organizers (cf. email addresses in the following section and in the authors' affiliation information).

\section{Important dates and information}
\begin{itemize}
\item May 15 - deadline for new benchmarks
\item June 1 - deadline for solver entrants
\item June 15 - deadline for final versions of solvers (but no new entrants after May 25).
\item July 17 - SMT workshop; completion of competition.
\item SMT-COMP mailing list: smt-comp@cs.nyu.edu
\item signup site for mailing list: \url{cs.nyu.edu/mailman/listinfo/smt-comp}
\item competition web site: \url{www.smtcomp.org}  or \url{smtcomp.sourceforge.net/2014}
\end{itemize}

@@

The annual Satisfiability Modulo Theories Competition~(SMT-COMP) is
held to spur advances in SMT solver implementations on benchmark
formulas of practical interest.  Public competitions are a well-known
means of stimulating advancement in software tools.  For example, in
automated reasoning, the CASC and SAT competitions for first-order and
propositional reasoning tools, respectively, have spurred significant
innovation in their fields~\cite{PSS02,leberre+03}.  More information
on the history and motivation for SMT-COMP can be found at the
SMT-COMP web site, \url{www.smtcomp.org}, and in reports on previous
competitions~(\cite{SMTCOMP-2008,BDOS08,SMTCOMP-FMSD,SMTCOMP-JAR,SMTCOMP2012,SMTEVAL2013}).
SMT-COMP 2014 is affiliated with the SMT workshop (\url {http://smt2014.it.uu.se/}) at the co-located meetings of CAV, IJCAR, and SAT at FLoC 2014 (\url{http://vsl2014.at/}); the competition is part of the
FLoC Olympic Games (\url{http://vsl2014.at/olympics/}).

@@

SMT-COMP 2014 will have two tracks: the conventional main track and a repeat of 2012's application track.
Within a track there are multiple divisions, where each division
uses benchmarks from a specific SMT-LIB logic (or group of logics).

The rest of this document, revised from the previous 
version,\footnote{Earlier versions of this document include contributions from
Clark Barrett, Albert Oliveras, Aaron Stump, Morgan Deters, David Cok, Alberto Griggio, and Roberto Bruttomesso.}
describes the rules and competition procedures for SMT-COMP~2014.
The principal changes from the previous competition rules are the following:
\begin{itemize}
\item The competition will be run with StarExec (\url{www.starexec.org}). Accordingly we are simplifying the competition somewhat, since some of the infrastructure needs to be recreated.
\item We will hold just a main and application track. [The demonstration unsat core, and proof generation tracks will be deferred until benchmarks and result analysis tools are reorganized.] We encourage submission of solvers and benchmarks for both tracks.
\item In the main track, we will measure comparative performance within all divisions to which at least two solvers are submitted. We will use the data to determine both sequential and parallel winners.

\item We will recognize winners as measured by number of benchmarks solved; we will also recognize best open source contributions and best new entrants, as described below. 
Also, importantly, the FLoC Olympic Games is sponsoring award ceremonies (with actual medals). A selection of the SMT-COMP winners (as described below) will be awarded FLoC Olympic Games medals.

\end{itemize}

All results of the competition will be made public. However, 
winners will be declared only for divisions in which at least two solvers are submitted.
Although the organizers may add other solvers for comparison purposes, only solvers that
are explicitly submitted by their authors are eligible to be designated as winners.

@@

\header{Solver format.} %
An entrant to SMT-COMP is an SMT solver
submitted by its authors using the StarExec (\url{http://www.starexec.org}) service.  
The execution service enables members of the
SMT research community to run solvers on jobs consisting of benchmarks
from the SMT-LIB benchmark library.  Jobs are run on a shared computer
cluster.  The execution service is provided free of
charge, but it does require registration to create a login account.  Registered users
may then upload their own solvers to run, or may run public solvers
already uploaded to the service.

For participation in SMT-COMP, a solver must be
uploaded to StarExec, made publicly available, and the organizers informed of its presence {\em and the logic divisions in which it is participating}. For clarity, the solver should have just one configuration, or at least, that the default configuration is the one to be used for competition.
A submission must also include a 32-bit unsigned
integer.  These numbers, collected from all submissions, are used
to seed the pseudo-random benchmark selection algorithm, as well
as the benchmark scrambler.

Information about how to configure and upload a solver is contained in the StarExec user guide on its wiki:
\url{https://wiki.uiowa.edu/display/stardev/User+Guide}.

\header{System description.} %
As part of a submission,
SMT-COMP entrants are encouraged to provide a short (1--2 pages) description of
the system.  This should include a list of all authors of the system,
 their present institutional affiliations, and any appropriate acknowledgements.  The programming
language(s) and basic SMT solving approach employed should be
described (\eg, lazy integration of a Nelson-Oppen combination with
SAT, translation to SAT, etc.).  System descriptions are encouraged to
include a URL for a web site for the submitted tool.  System descriptions may be submitted after the solver deadline, but to be useful should be sent to the organizers before the competition ends (July 16).

\header{Other systems.} %
The organizers' intent is to promote as wide a comparison among solver options as possible.
However, if the number of solver submissions is too large for the
computational resources available to the competition, the organizers reserve the right not to
accept multiple versions of solvers from the same solver team.  The organizers also reserve the right to
include other systems of interest (such as entrants in previous competitions) in the
competition.

@@

\header{Attendance.} %
As with previous SMT-COMPs, submitters of an SMT-COMP entrant need not 
be physically present at the competition to
participate or win.

\subsection*{Deadlines} %

%%%%%% Make main and parallel competitions together.

\paragraph{Main competition track.} %
SMT-COMP entries must be submitted via StarExec and email to the organizers by the end of June~1, 2014 (anywhere on earth).  At that time new entrants will be
closed to the public to prepare for the competition, with the
exception that resubmissions of existing entries will be accepted
until the end of June~15, 2014 (anywhere on earth).  We strongly encourage
participants to use this grace period \emph{only} for the
purpose of fixing any bugs that may be discovered and not for adding
new features as there will be no opportunity to do extensive testing
using StarExec after the initial deadline.

Competition solvers should be submitted with just one StarExec configuration so that it is
clear what configuration should be used for the competition. Separately named StarExec solvers can wrap the same underlying tool with a different configuration, if needed, for the application track.

The versions that are present on the execution service at the conclusion of the
grace period will be the ones used for the competition, and versions
submitted after this time will not be used.  The organizers reserve
the right to start the competition itself at any time after the open
of the New York Stock Exchange on the day after the final solver deadline.  See
Section~\ref{sec:timeline} below for a full timeline.

\paragraph{Application track.} %
The same deadlines and procedure for submitting to the main track will be used
also for the application track. Submissions to both the application track and
the main competition are independent: participants must submit explicitly to
both events to participate in both, and they may submit different (or
differently configured) solvers to each.  Benchmarks will be scrambled also
for this division, using the same scrambler and seed as the main track.
Entrants should still include a system description, as for the regular
competition.

The infrastructure for the application track is in the process of being recreated for StarExec.
If it is not ready in time, the organizers may choose to drop this track or only run it as a demonstration track.

@@

Solvers will be publicly evaluated in the main and application tracks, each in several logic divisions. 
All divisions with at least two solver entrants are competition divisions, i.e., a winner will be announced. The other divisions are exhibition divisions, i.e., solvers are evaluated and results posted, but no awards are announced.

@@

\header{Dates of competition.}
%
We anticipate that the bulk of the main track of the competition will take place during
the weeks leading up to SMT 2014, from about June~16 to July~17.  Results will be
announced at SMT 2014 and in an awards ceremony scheduled by the FLoC Olympic Games coordinator.
  Intermediate results
will be regularly posted to the SMT-COMP website as the competition
runs.

The organizers will prioritize the running of the competition tracks, and may shift the
time period or order of the competition or exhibition tracks in order to complete SMT-COMP
prior to the SMT workshop and awards ceremony.

\header{Input and Output.}
%
A participating solver must read a single benchmark script presented as its first argument.
Note that previous competitions presented benchmarks to a solver on its standard input
channel, and that the application track will send command inputs to the solver's standard input.
Accordingly, solvers are advised to support both input mechanisms. The organizers have easily adapted solvers that use only the standard input to StarExec with a simple wrapper script.

 The script is in the concrete syntax of the SMT-LIB format,
version 2.0, though with a restricted set of commands.   A
\textbf{benchmark script} is just a text file containing a sequence of SMTLIBv2 commands in which:

\begin{enumerate}
\item The (single) $\akey{set-logic}$ command setting the benchmark's
logic is the first command after any $\akey{set-option}$ commands described below.
\item In the main track, there may be a (single) \akey{set-option :print-success ...} command. Note that {\tt success} outputs are ignored by the post-processor used by the competition; in 
the application track, the :print-success option may not be disabled.
\item The application track executor will send an initial $\akey{set-option :print-success true}$ command to the solver. 
\item The script next optionally contains (any number of) $\akey{set-info}$ commands, which may be ignored by the solver. Benchmarks are expected to have the following \akey{set-info} commands: \akey{(set-info :smt-lib-version 2.0)}, \akey{(set-info :category ...)}, \akey{(set-info :source ...)} (indicating the authorship or tool that produced the benchmark), and \akey{(set-info :status ...)}.
\item The $\akey{exit}$ command is the last command.
\item For tracks other than the application track, there is exactly one $\akey{check-sat}$ command,
following possibly several $\akey{assert}$ commands.
\item For the application track, there are one or more \akey{check-sat} commands, 
  each preceded by one or more \akey{assert} commands; 
  there may also be zero or more \akey{push 1} commands, 
  and zero or more \akey{pop 1} commands, consistent with the uses of those commands in
  the SMT-LIB standard. Each \akey{check-sat} command may be preceded by a \akey{set-info :status} command that indicates the expected result of the 
  \akey{check-sat} command.

\item The formulas in the script belong to the benchmark's logic, with
any free symbols declared in the script.
\item All sorts declared with a $\akey{declare-sort}$ command must have zero arity.

\item Application track scripts may have $\akey{define-sort}$ and $\akey{define-sort}$ commands. 
\item No other commands besides the ones just mentioned may be used.
\item Named formulas are not used in benchmark scripts.

\noindent The SMT-LIB format specification is publicly
available from the ``Documents'' section of the SMT-LIB
website~\cite{SMT-LIB}.  Solvers will be given formulas just from the
Problem Divisions indicated during their submission to SMT-COMP.

@@

The main track competition will consist of selected benchmarks in each of the
logic divisions.
Each benchmark script will be presented to the solver as its first command-line argument.
Each SMT-COMP entrant is then expected to attempt to report on its
standard output channel whether the formula is satisfiable
(``\texttt{sat}'', in lowercase, without the quotation marks) or unsatisfiable
(``\texttt{unsat}'').  An entrant may also report ``\texttt{unknown}''
to indicate that it cannot determine satisfiability of the formula.

@@

\header{Timeouts.} %
Each SMT-COMP solver will be executed on an
unloaded competition machine for each given benchmark, up to a fixed
time limit.  The time limit is yet to be determined, but it is
anticipated to be 1500 seconds (25 minutes) of CPU time and the same limit on wall clock time.\footnote{The time limit may be adjusted once we understand the resources and capabilities of the StarExec service and the number of competition solvers.} 
Solvers that take more than this
time limit will be killed.  Solvers are allowed to spawn other
processes.  These will be killed at approximately the same time as the
first started process.

The StarExec service also limits the memory consumption of the solver processes. We expect the limit per solver-benchmark job-pair to be quite high, on the order of 100GB. The values of both the timeout limit and the memory limit are available to a solver process through environment variables defined and set by the StarExec service (see the StarExec user guide for more information).

\header{Aborts and unparsable output.} %
Solvers which exit before the time
limit without reporting a result (\ie\ due to exhausting memory, crashing,
or producing output other than \texttt{sat}, \texttt{unsat}, or
\texttt{unknown})
will be considered to have aborted. 
Also, as a further measure to prevent misjudgments of solvers,
any  ``\texttt{success}'' outputs will be 
ignored.\footnote{
Note that SMT-LIBv2 requires that a solver produce a ``\texttt{success}'' answer
after each \akey{set-logic}, \akey{declare-sort}, \akey{declare-fun} and
\akey{assert} command (among others), unless the option
\akey{:print-success} is set to false; ignoring the
\texttt{success} outputs therefore allows for submitting fully-compliant
solvers without the need of a wrapper script, while still allowing entrants
of previous competitions to run without changes.}

@@

The solver is executed with a temporary directory as the current working directory.
Any generated files should be produced there, not, for example, in the system tmp area.
The temporary directory and its contents are deleted after the job is complete.

@@

The application track evaluates SMT solvers when interacting
with an external verification framework, \eg, a model
checker. This interaction, ideally, happens by means of an online
communication between the model checker and the solver: the model
checker repeatedly sends queries to the SMT solver, which in turn
answers either \texttt{sat} or \texttt{unsat}.  In this interaction an SMT-solver is
required to accept queries incrementally via its {\em standard input channel}.

In order to facilitate the evaluation of the solvers in this track, we
will set up a ``simulation'' of the aforementioned interaction, as was done in 2012. In
particular each benchmark in the application track represents a realistic
communication trace, containing multiple \akey{check-sat} commands (possibly
with corresponding \akey{push 1}/\akey{pop 1} commands), which
is parsed by a {\em trace executor}. The trace executor serves the following purposes:
\begin{itemize}
\item it simulates the online interaction by sending single queries to the SMT solver
      (through stdin);
\item it prevents ``look-ahead'' behaviors of SMT solvers;
\item it records time and answers for each call, possibly aborting the execution
      in case of a wrong answer;
\item it guarantees a fair execution for all solvers by abstracting from any possible
      crash, misbehavior, etc. that may happen on the model checker side.
\end{itemize}
Note that the executor terminates processing the script upon receiving an incorrect response.

\header{Input and Output.}
Participating solvers will be connected to a trace executor 
which will incrementally send commands to the standard input channel of the solver
and read responses from the standard output channel of the solver.
The commands will be taken from an \textbf{incremental benchmark script},
which is an SMT-LIBv2 script which satisfies the rules for an application script given above.
Note also that the trace executor will send a single 
\akey{set-option :print-success true} command to the solver before 
sending commands from the incremental benchmark script.

\medskip
Solvers must respond to each command sent by the trace executor, 
with the answers defined in the SMT-LIB 2.0 format specification, that is,
with a \texttt{success} answer for 
\akey{set-option}, 
\akey{set-logic}, 
\akey{declare-sort}, 
\akey{declare-fun}, 
\akey{define-sort}, 
\akey{define-fun}, 
\akey{assert}, 
\akey{push 1}, 
and \akey{pop 1} 
commands,
 with a \texttt{sat}, \texttt{unsat}, or \texttt{unknown} 
for \akey{check-sat} commands. There will be no \akey{get-unsat-core}, \akey{get-proof}, \akey{get-model}, \akey{get-assignments}, or \akey{get-values} commands in these benchmarks.

\header{Timeouts.}
A time limit is set for the whole execution of each application
benchmark (consisting of multiple queries).  We anticipate the
timeout to be around 25~minutes (as it has been in the past).\footnote{The time limit may be adjusted once we understand the resources and capabilities of the StarExec service and the number of competition solvers.}

@@

In 2012 the competition ran only some logics as competitive divisions.
In 2014 we will revert to previous practice and run each logic as a competitive division, if there are two or more solvers submitted.

@@

\header{Benchmark sources.} %
Benchmark formulas for each division of the main track
will be drawn from the SMT-LIB library. The deadline for new benchmarks is
May 15. The organizers will be checking and curating these until June 1;
the organizers reserve the option to exclude new benchmarks if any prove problematic for some reason.
SMT-COMP
attempts to give preference to benchmarks that are ``real-world,'' in
the sense of coming from or having some intended application outside
SMT.

\header{Benchmark availability.} % 
The deadline for new benchmarks is May 15, 2014.
The benchmarks will be made available as soon as possible after that, as they are checked and curated.
The goal for completing the set of benchmarks is June 1.
The set of benchmarks selected for the competition will be published when the
competition begins.

@@

Each benchmark in SMT-LIB also has a \emph{category}.  There are four possible
categories:
%
\begin{itemize}
\item \emph{check.} These benchmarks are hand-crafted to test whether
  solvers support specific features of each division.  In particular,
  there are checks for integer completeness (\ie\ benchmarks that are
  satisfiable under the reals but not under the integers) and big
  number support (\ie\ benchmarks that are likely to fail if integers
  cannot be represented beyond some maximum value, such as
  $(2^{31}-1)$).

>>>>

\documentclass[12pt]{article}

\usepackage{color}
\usepackage{times}
\usepackage{fullpage}
\usepackage{hyperref}

\usepackage{draftwatermark}

@@

\newcommand{\remark}[1]{{\small\textcolor{red}{[#1]}}}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

\title{10th International Satisfiability Modulo Theories Competition
  (SMT-COMP 2015): Rules and Procedures}

@@

\author{%
Sylvain Conchon \\
Paris-Sud University \\
France \\
{\small\href{mailto:Sylvain.Conchon@lri.fr}{\tt Sylvain.Conchon@lri.fr}} \\
\and
David D{\'e}harbe \\
Federal University of Rio Grande do Norte \\
Brazil \\
{\small\href{mailto:David.Deharbe@loria.fr}{\tt David.Deharbe@loria.fr}} \\
\and
Tjark Weber \\
Uppsala University \\
Sweden \\
{\small\href{mailto:tjark.weber@it.uu.se}{\tt tjark.weber@it.uu.se}} \\
\\
}

@@

\noindent Comments on this document should be emailed to the SMT-COMP
mailing list (see below) or, if necessary, directly to the organizers.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Communication}
\label{sec:communication}

Interested parties should subscribe to the SMT-COMP mailing list.
Important late-breaking news and any necessary clarifications and
edits to these rules will be announced there, and it is the primary
way that such announcements will be communicated.

\begin{itemize}
\item SMT-COMP mailing list: \href{mailto:smt-comp@cs.nyu.edu}{\tt
  smt-comp@cs.nyu.edu}
\item Sign-up site for the mailing list:
  \url{cs.nyu.edu/mailman/listinfo/smt-comp}
\end{itemize}

\noindent Additional material will be made available at the
competition web site, \url{www.smtcomp.org} or
\url{smtcomp.sourceforge.net/2015}.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\newpage

\section{Important Dates}
\label{sec:important}

\begin{description}
\item[May~1] Deadline for new benchmark contributions.
\item[June~1] Deadline for first versions of solvers (for all tracks),
  including information about which tracks and divisions are being
  entered, and magic numbers for benchmark scrambling.
\item[June~7] Final versions of competition tools (e.g., benchmark
  scrambler) are made available.  Benchmark libraries are frozen.
\item[June~14] Deadline for final versions of solvers, including
  system descriptions.
\item[June~15] Opening value of NYSE Composite Index used to complete
  random seed for benchmark scrambling.
\item[July~18/19] SMT Workshop; end of competition, presentation of
  results.
\end{description}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

The annual Satisfiability Modulo Theories Competition~(SMT-COMP) is
held to spur advances in SMT solver implementations on benchmark
formulas of practical interest.  Public competitions are a well-known
means of stimulating advancement in software tools.  For example, in
automated reasoning, the CASC and SAT competitions for first-order and
propositional reasoning tools, respectively, have spurred significant
innovation in their fields~\cite{leberre+03,PSS02}.  More information
on the history and motivation for SMT-COMP can be found at the
competition web site, \url{www.smtcomp.org}, and in reports on
previous
competitions~(\cite{SMTCOMP2012,SMTEVAL2013,SMTCOMP-JAR,SMTCOMP-FMSD,SMTCOMP-2008,BDOS08}).

SMT-COMP~2015 is affiliated with the SMT Workshop~2015
(\url{http://smt2015.csl.sri.com/}), which is associated with CAV~2015
(\url{http://i-cav.org/2015/}).  The SMT Workshop and the main CAV
conference will include a block of time to present the competitors and
results of the competition.

@@

SMT-COMP 2015 will have two tracks: the conventional main track and an
application (i.e., incremental) track.  Within each track there are
multiple divisions, where each division uses benchmarks from a
specific SMT-LIB logic (or group of logics). \remark{Additional
  tracks, e.g., for unsat-core extraction or proof generation, are
  currently being considered.}  We will recognize winners as measured
by number of benchmarks solved; we will also recognize solvers based
on additional criteria.

The rest of this document, revised from the previous
version,\footnote{Earlier versions of this document include
  contributions from Clark Barrett, Roberto Bruttomesso, David Cok,
  David D{\'e}harbe, Morgan Deters, Alberto Griggio, Albert Oliveras,
  Aaron Stump, and Tjark Weber.}  describes the rules and competition
procedures for SMT-COMP~2015.  The principal changes from the previous
competition rules are the following:
\begin{itemize}
\item All solvers will be run with $n=4$ cores available, and the CPU
  timeout will be $n$~times the wall-clock timeout.  We will recognize
  both best sequential and best parallel performance in all divisions.
\item In addition to recognizing the best solver in each division, we
  will recognize the three solvers that perform best according to
  competition-wide criteria, emphasizing breadth of supported logics.
  These criteria will be a refinement of the criteria used for the
  FLoC Olympic Games medals in~2014.
\item \remark{Pending availability of suitable benchmarks and solvers,
  we may include a new division for floating-point arithmetic.  This
  should be considered experimental in 2015, and will not be included
  in the competition-wide ranking.}
\item \remark{Benchmark selection and relative weight of benchmark
  families are under review.}
\end{itemize}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

\header{Solver submission.}
%
An entrant to SMT-COMP is an SMT solver submitted by its authors using
the StarExec (\url{http://www.starexec.org}) service.  The execution
service enables members of the SMT research community to run solvers
on jobs consisting of benchmarks from the SMT-LIB benchmark library.
Jobs are run on a shared computer cluster.  The execution service is
provided free of charge, but it does require registration to create a
login account.  Registered users may then upload their own solvers to
run, or may run public solvers already uploaded to the service.

For participation in SMT-COMP, a solver must be uploaded to StarExec,
made publicly available, and the organizers informed of its presence
{\em and the tracks and divisions in which it is
  participating}. StarExec supports solver configurations; for
clarity, each submitted solver must have one configuration only.  A
submission must also include a 32-bit unsigned integer.  These
numbers, collected from all submissions, are used to seed the
pseudo-random benchmark selection algorithm, as well as the benchmark
scrambler.

Information about how to configure and upload a solver is contained in
the StarExec user guide,
\url{https://wiki.uiowa.edu/display/stardev/User+Guide}.

\header{System description.}
%
As part of a submission, SMT-COMP entrants are encouraged to provide a
short (1--2 pages) description of the system.  This should include a
list of all authors of the system, their present institutional
affiliations, and any appropriate acknowledgements.  The programming
language(s) and basic SMT solving approach employed should be
described (e.g., lazy integration of a Nelson-Oppen combination with
SAT, translation to SAT, etc.).  System descriptions are encouraged to
include a URL for a web site for the submitted tool.  System
descriptions may be submitted after the solver deadline, but to be
useful should be sent to the organizers before the competition ends.

\header{Other systems.}
%
The organizers' intent is to promote as wide a comparison among
solvers and solver options as possible.  However, if the number of
solver submissions is too large for the computational resources
available to the competition, the organizers reserve the right not to
accept multiple versions of solvers from the same solver team.  The
organizers also reserve the right to include other systems of interest
(such as entrants in previous competitions) in the competition.

@@

\header{Attendance.}
%
Submitters of an SMT-COMP entrant need not be physically present at
the competition to participate or win.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\subsection*{Deadlines}

SMT-COMP entrants must be submitted via StarExec (solvers) \emph{and}
email to the organizers (accompanying information) until the end of
{\bf June~1, 2015} anywhere on earth.  After this date no new entrants
will be accepted.  However, updates to existing entrants on StarExec
will be accepted until the end of {\bf June~14, 2015} anywhere on
earth.

We strongly encourage participants to use this grace period
\emph{only} for the purpose of fixing any bugs that may be discovered,
and not for adding new features, as there may be no opportunity to do
extensive testing using StarExec after the initial deadline.

The solver versions that are present on StarExec at the conclusion of
the grace period will be the ones used for the competition, and
versions submitted after this time will not be used.  The organizers
reserve the right to start the competition itself at any time after
the open of the New York Stock Exchange on the day after the final
solver deadline.

These deadlines and procedures apply equally to all tracks of the
competition.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

Solvers will be publicly evaluated in all tracks and divisions into
which they have been entered.  All results of the competition will be
made public.

@@

\header{Dates of competition.}
%
The bulk of the computation will take place during the weeks leading
up to SMT~2015, from about June~15 to July~17.  Intermediate results
will be regularly posted to the SMT-COMP website as the competition
runs.

The organizers reserve the right to prioritize certain competition
tracks or divisions to ensure their timely completion, and in
exceptional circumstances to complete divisions after the SMT
Workshop.

\header{Input and output.}
%
In the main track, a participating solver must read a single benchmark
script, whose name is presented as the solver's first argument.  In
the application track, a trace executor will send commands from a
benchmark script to the solver's standard input channel.

The benchmark script is in the concrete syntax of the SMT-LIB format,
version~2.0, though with a restricted set of commands.  A benchmark
script is a text file containing a sequence of SMT-LIB~v2 commands in
which:
%
\begin{enumerate}
\item The (single) $\akey{set-logic}$ command setting the benchmark's
  logic is the first command after any $\akey{set-option}$ commands
  described below.
\item In the main track, there may be a (single) \akey{set-option
  :print-success ...} command. Note that {\tt success} outputs are
  ignored by the post-processor used by the competition; in the
  application track, the :print-success option may not be disabled.
\item The application track executor will send an initial
  $\akey{set-option :print-success true}$ command to the solver.
\item The script next optionally contains (any number of)
  $\akey{set-info}$ commands, which may be ignored by the
  solver. Benchmarks are expected to have the following
  \akey{set-info} commands: \akey{(set-info :smt-lib-version 2.0)},
  \akey{(set-info :category ...)}, \akey{(set-info :source ...)}
  (indicating the authorship or tool that produced the benchmark), and
  \akey{(set-info :status ...)}.
\item The $\akey{exit}$ command is the last command.
\item For tracks other than the application track, there is exactly
  one $\akey{check-sat}$ command, following possibly several
  $\akey{assert}$ commands.
\item For the application track, there are one or more
  \akey{check-sat} commands, each preceded by one or more
  \akey{assert} commands; there may also be zero or more \akey{push 1}
  commands, and zero or more \akey{pop 1} commands, consistent with
  the uses of those commands in the SMT-LIB standard. Each
  \akey{check-sat} command may be preceded by a \akey{set-info
    :status} command that indicates the expected result of the
  \akey{check-sat} command.
\item The formulas in the script belong to the benchmark's logic, with
  any free symbols declared in the script.
\item All sorts declared with a $\akey{declare-sort}$ command must
  have zero arity.
\item Application track scripts may have $\akey{define-sort}$
  commands.
\item No other commands besides the ones just mentioned may be used.
\item Named formulas are not used in benchmark scripts.
\end{enumerate}
%
The SMT-LIB format specification is publicly available from the
``Documents'' section of the SMT-LIB website~\cite{SMT-LIB}.  Solvers
will be given formulas only from the divisions into which they have
been entered.

\header{Timeouts.}
%
Each SMT-COMP solver will be executed on a dedicated processor with
4~cores of a competition machine for each given benchmark, up to a
fixed time limit.  Detailed machine specifications are available on
the competition web site.  The time limit is yet to be determined, but
it is anticipated to be 40~minutes of wall-clock time (and 160~minutes
of CPU time).\footnote{The time limit may be adjusted once we know the
  number of competition entrants and eligible benchmarks.}  Solvers
that take more than this time limit will be killed.  Solvers are
allowed to spawn other processes; these will be killed at
approximately the same time as the first started process.

The StarExec service also limits the memory consumption of the solver
processes.  We expect the limit per solver-benchmark job-pair to be on
the order of 60\,GB.  The values of both the timeout limit and the
memory limit are available to a solver process through environment
variables defined and set by the StarExec service (see the StarExec
user guide for more information).

@@

The main track competition will consist of selected benchmarks in each
of the logic divisions.  Each benchmark script will be presented to
the solver as its first command-line argument.  Each SMT-COMP entrant
is then expected to attempt to report on its standard output channel
whether the formula is satisfiable (\texttt{sat}, in lowercase) or
unsatisfiable (\texttt{unsat}).  An entrant may also report
\texttt{unknown} to indicate that it cannot determine satisfiability
of the formula.

@@

\header{Aborts and unparsable output.}
%
Solvers which exit before the time limit without reporting a result
(i.e., due to exhausting memory or crashing) and do not produce output
that includes \texttt{sat}, \texttt{unsat} or \texttt{unknown} will be
considered to have aborted.  Also, as a further measure to prevent
misjudgments of solvers, any \texttt{success} outputs will be
ignored.\footnote{ Note that SMT-LIB~v2 requires that a solver produce
  a \texttt{success} answer after each \akey{set-logic},
  \akey{declare-sort}, \akey{declare-fun} and \akey{assert} command
  (among others), unless the option \akey{:print-success} is set to
  false.  Ignoring the \texttt{success} outputs therefore allows for
  submitting fully-compliant solvers without the need of a wrapper
  script, while still allowing entrants of previous competitions to
  run without changes.}

@@

The solver is executed with a temporary directory as the current
working directory.  Any generated files should be produced there (and
not, say, in the system's {\tt /tmp} directory).  The StarExec system
sets a limit on the amount of disk storage permitted---typically
20\,GB (see the User Guide).  The temporary directory and its contents
are deleted after the job is complete.

@@

The application track evaluates SMT solvers when interacting with an
external verification framework, e.g., a model checker. This
interaction, ideally, happens by means of an online communication
between the model checker and the solver: the model checker repeatedly
sends queries to the SMT solver, which in turn answers either
\texttt{sat} or \texttt{unsat}.  In this interaction an SMT-solver is
required to accept queries incrementally via its {\em standard input
  channel}.

In order to facilitate the evaluation of the solvers in this track, we
will set up a ``simulation'' of the aforementioned interaction. In
particular each benchmark in the application track represents a
realistic communication trace, containing multiple \akey{check-sat}
commands (possibly with corresponding \akey{push 1}/\akey{pop 1}
commands), which is parsed by a {\em trace executor}. The trace
executor serves the following purposes:
\begin{itemize}
\item it simulates the online interaction by sending single queries to
  the SMT solver (through stdin);
\item it prevents ``look-ahead'' behaviors of SMT solvers;
\item it records time and answers for each call, possibly aborting the
  execution in case of a wrong answer;
\item it guarantees a fair execution for all solvers by abstracting
  from any possible crash, misbehavior, etc. that may happen on the
  model checker side.
\end{itemize}
Note that the executor terminates processing the script upon receiving
an incorrect response.

The disk space and memory limits on the application track are the same
as for the main track.

\header{Input and Output.}
%
Participating solvers will be connected to a trace executor which will
incrementally send commands to the standard input channel of the
solver and read responses from the standard output channel of the
solver.  The commands will be taken from an \textbf{incremental
  benchmark script}, which is an SMT-LIBv2 script which satisfies the
rules for an application script given above.  Note also that the trace
executor will send a single \akey{set-option :print-success true}
command to the solver before sending commands from the incremental
benchmark script.

Solvers must respond to each command sent by the trace executor, with
the answers defined in the SMT-LIB 2.0 format specification, that is,
with a \texttt{success} answer for \akey{set-option},
\akey{set-logic}, \akey{declare-sort}, \akey{declare-fun},
\akey{define-sort}, \akey{define-fun}, \akey{assert}, \akey{push 1},
and \akey{pop 1} commands, and with a \texttt{sat}, \texttt{unsat}, or
\texttt{unknown} for \akey{check-sat} commands. There will be no
\akey{get-unsat-core}, \akey{get-proof}, \akey{get-model},
\akey{get-assignments}, or \akey{get-values} commands in these
benchmarks.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

@@

\header{Benchmark sources.}
%
Benchmark formulas for each division of the main track will be drawn
from the SMT-LIB library. The deadline for new benchmarks is {\bf May
  1, 2015}. The organizers will be checking and curating these until
{\bf June 7, 2015}; the organizers reserve the option to exclude new
benchmarks if any prove problematic for some reason.  SMT-COMP
attempts to give preference to benchmarks that are ``real-world,'' in
the sense of coming from or having some intended application outside
SMT.

\header{Benchmark availability.}
%
New benchmarks will be made available as soon as possible after the
benchmark submission deadline, as they are checked and curated.  The
set of benchmarks selected for the competition will be published when
the competition begins.

@@

Each benchmark in SMT-LIB also has a \emph{category}.  There are four possible
categories:
%
\begin{itemize}
\item \emph{check.} These benchmarks are hand-crafted to test whether
  solvers support specific features of each division.  In particular,
  there are checks for integer completeness (i.e., benchmarks that are
  satisfiable under the reals but not under the integers) and big
  number support (i.e., benchmarks that are likely to fail if integers
  cannot be represented beyond some maximum value, such as
  $(2^{31}-1)$).

\item \emph{industrial.} These benchmarks come from some real
  application and are produced by tools such as bounded model
  checkers, static analyzers, extended static checkers, etc.

\item \emph{random.} These benchmarks are randomly generated.

\item \emph{crafted.} This category is for all other benchmarks.
  Usually, benchmarks in this category are designed to be particularly
  difficult or to test a specific feature of the logic.
\end{itemize}

\header{Benchmark selection.} \remark{This section is under review.}
%
Each benchmark is assigned a difficulty. For 2015, the difficulty
measure is the CPU time taken by the best performing solver in the
2014 SMT-COMP.  For new benchmarks, we will determine an approximate
difficulty using trial runs.  The difficulty values will be posted as
soon after May~1 as possible.  The difficulty values are best
estimates by the organizers and may not always reflect actual
difficulty.
